A Reduced Order Model for Porous Media Flows: Coupling Linear Regression with Principal Component Analysis (PCA) using High-Order Singular Value Decomposition (HOSVD) in Upscaling Calculations for Permeability in Reservoir Simulations

### 2018-12-03 Papers/AppliedMaths2018/Resubmission/Stochastic_Upscaling_UK_AMM_2018b.tex
files to be modified in the next 250min

\subsubsection{BaseCase Pre-processing Algorithm}\label{subsubsection:basecase_preprocess_algorithm}
\begin{enumerate}[1.]
  \item A function is used to randomly generate uniform (\ie Gaussian) dataset for each block. The range for each block is as below:
  \begin{enumerate}[a)]
    \item \textbf{K1} : $700 - 900$ mD
    \item \textbf{K2} : $50 - 150$ mD
    \item \textbf{K3} : $100 - 200$ mD
    \item \textbf{K4} : $1000 - 1200$ mD
  \end{enumerate}                                                    
  \item The number of data points generated for each block is $1600$ and the data values are randomly allocated within each block in a structured pattern. Albeit, the simulation mesh is unstructured.
  \item Merge the dataset from different blocks into a single file retaining the original ``structured pattern'' (in step 1 above) for the randomly positioned dataset for each block. This file will be used by the CVFEM-based flow simulator to prescribe the permeability field for the base case.
\end{enumerate}

\subsubsection{PDFCase Pre-processing Algorithm}\label{subsubsection:pdfcase_preprocess_algorithm}
\begin{enumerate}[1.]
  \item The datasets from step one of the pre-processing algorithm for the base case (Section~\ref{subsubsection:basecase_preprocess_algorithm}) are used to calculate mean and standard deviation for each block of the permeability field.
  \item A function which generates a Gaussian dataset using arithmetic mean and standard deviation is used to stochastically produce a quarter of the data points from the BaseCase for each of the four blocks  (\ie $400$ data points for each block).
  \item The data values in the dataset for each block is randomly positioned within each block in a structured pattern. As in the base case, the simulation grid is unstructured.
  \item Finally, the structured/arranged dataset from each block is combined into a single file and then used by the flow simulator, which merge the permeability mapping to the mesh with the appropriate element-pairs.
\end{enumerate}

\subsubsection{2-D PCA (\ie{using SVD}) \& LinReg Case Pre-processing Algorithm}\label{subsubsection:svdcase_preprocess_algorithm}
\begin{enumerate}[1.]
  \item The permeability field obtained in step 3 of the pre-processing algorithm for the base case is factorized using SVD to obtain $\mathbf{U}, \Sigma,$ and $\mathbf{V}$.
  \item A selected number ($z$) of singular values are retained in decreasing order of magnitude to form a new diagonal matrix, $\Sigma_{new}$. (This step is optional, it is the conventional PCA dimensionality reduction)
  \item The first $z$ columns of $\mathbf{U}$ and $\mathbf{V}$ are selected to form new truncated matrices call $\mathbf{U_{new}}$ and $\mathbf{V_{new}}$ respectively. (This step is necessary only if the optional step in 2 above is executed)
  \item Within each column of $\mathbf{U_{new}}$ and $\mathbf{V_{new}}$, a linear interpolation is performed to form a new column with the number of data points in each column reduced to half of the original column. These new columns are recombined to form $\mathbf{U_{new2}}$ and $\mathbf{V_{new2}}$.
  \item The permeability field used for the SVD case is the matrix $\mathbf{A_{new2}}$, which is defined as the product of $\mathbf{U_{new2}} \mathbf{\Sigma_{new}} \mathbf{V_{new2}^{\intercal}}$. It is a permeability field with a quarter of the data values in the base case and therefore a permeability field with reduced order.  
\end{enumerate}

>>>>

\subsubsection{BaseCase Pre-processing Algorithm}\label{subsubsection:basecase_preprocess_algorithm}
\begin{description}
  \item[Step 1:] A function is used to randomly generate uniform (\ie Gaussian) dataset for each block. The range for each block is as below:
  \begin{enumerate}[a)]
    \item \textbf{K1} : $700 - 900$ mD;
    \item \textbf{K2} : $50 - 150$ mD;
    \item \textbf{K3} : $100 - 200$ mD;
    \item \textbf{K4} : $1000 - 1200$ mD;
  \end{enumerate}                                                    
  \item[Step 2:] The number of data points generated for each block is $1600$, and data is randomly distributed within each block in a structured pattern. Albeit, the mesh is unstructured;
  \item[Step 3:] Merge the dataset from different blocks into a single file retaining the original ``structured pattern'' ({\it Step 1}) for the randomly positioned dataset for each block. This data-structure will be used by the CVFEM-based flow simulator to prescribe the permeability dataset for the BaseCase configuration.
\end{description}

\subsubsection{PDFCase Pre-processing Algorithm}\label{subsubsection:pdfcase_preprocess_algorithm}
\begin{description}
  \item[Step 1:] Datasets from {\it Step 1} of Section~\ref{subsubsection:basecase_preprocess_algorithm} are used to calculate mean and standard deviation for each block of permeability data;
  \item[Step 2:] A function which generates a Gaussian dataset using arithmetic mean and standard deviation is used to stochastically produce a quarter of the data points from the BaseCase configuration for each of the four blocks (\ie $400$ data points for each block);
  \item[Step 3:] Data values in the dataset for each block is randomly positioned within each block in a structured pattern. Similarly to the BaseCase configuration, simulation mesh grid is also unstructured;
  \item[Step 4:] Finally, the structured/arranged dataset from each block is combined into a single data-structure and then used by the flow simulator, which merge the permeability mapping to the mesh with the appropriate element-pairs (in simulations shown here, \PN[1]{2} and \PNDG[1]{1} element-pairs are used).
\end{description}

\subsubsection{2-D PCA (SVD) \& LinReg Case Pre-processing Algorithm}\label{subsubsection:svdcase_preprocess_algorithm}
\begin{description}
  \item[Step 1:] Permeability data obtained in {\it Step 3}  of Section~\ref{subsubsection:basecase_preprocess_algorithm} is factorized using SVD method to obtain $\mathbf{U}, \Sigma,$ and $\mathbf{V}$;
  \item[Step 2:] A selected number ($z$) of singular values are retained in decreasing order of magnitude to form a new diagonal matrix, $\Sigma_{new}$ (this step is optional as it is the conventional PCA dimensionality reduction);
  \item[Step 3:] First $z$ columns of $\mathbf{U}$ and $\mathbf{V}$ are selected to form new truncated matrices called $\mathbf{U_{new}}$ and $\mathbf{V_{new}}$, respectively. This step is necessary only if the optional step in 2 above is executed;
  \item[Step 4:] Within each column of $\mathbf{U_{new}}$ and $\mathbf{V_{new}}$, a linear interpolation is performed to form a new column with the number of data points in each column reduced to half of the original column. These new columns are recombined to form $\mathbf{U_{new2}}$ and $\mathbf{V_{new2}}$;
  \item[Step 5:] Permeability dataset used for the PCA-LinReg Case is the matrix $\mathbf{A_{new2}}$, which is defined as the product of $\mathbf{U_{new2}} \mathbf{\Sigma_{new}} \mathbf{V_{new2}^{\intercal}}$. It is a permeability dataset with a quarter of the data values of the BaseCase, \ie a permeability dataset with reduced order.  
\end{description}
